### 2025-07-29 articles/arxiv/tables/side_by_side_similarity.tex
move figures and tables to arxiv folder (#7)

\begin{tabular}{lr|lr|lr|lr}

      \multicolumn{2}{c|}{Cam Ward} &
      \multicolumn{2}{c|}{Dillon Gabriel} &
      \multicolumn{2}{c|}{Shedeur Sanders} &
      \multicolumn{2}{c}{Jaxson Dart}
    \\
Comp & Score & Comp & Score & Comp & Score & Comp & Score \\
% latex table generated in R 4.5.0 by xtable 1.8-4 package
% Mon Jul 28 15:03:22 2025
  \hline
Johnny Manziel & 2.3\% & Mason Rudolph & 2.0\% & Mason Rudolph & 1.9\% & Teddy Bridgewater & 1.9\% \\ 
  Marcus Mariota & 2.3\% & Dwayne Haskins & 1.9\% & Kenny Pickett & 1.9\% & Tajh Boyd & 1.6\% \\ 
  Baker Mayfield & 2.2\% & Kenny Pickett & 1.9\% & Philip Rivers & 1.9\% & Russell Wilson & 1.5\% \\ 
  Philip Rivers & 2.1\% & C.J. Stroud & 1.9\% & Ben Roethlisberger & 1.8\% & John Beck & 1.5\% \\ 
  Trevor Lawrence & 2.0\% & Andrew Luck & 1.9\% & Case Keenum & 1.8\% & Drake Maye & 1.4\% \\ 
  Matt Leinart & 2.0\% & Bo Nix & 1.9\% & Russell Wilson & 1.8\% & Zach Terrell & 1.4\% \\ 
  Russell Wilson & 2.0\% & Trevor Lawrence & 1.9\% & Trevone Boykin & 1.8\% & Kevin Hogan & 1.2\% \\ 
  Lamar Jackson & 1.9\% & Philip Rivers & 1.8\% & Kellen Moore & 1.7\% & Blake Bortles & 1.2\% \\ 
  Deshaun Watson & 1.9\% & Aaron Rodgers & 1.7\% & Andrew Luck & 1.7\% & Sam Howell & 1.2\% \\ 
  Case Keenum & 1.9\% & Bryce Young & 1.6\% & C.J. Stroud & 1.7\% & Patrick Mahomes & 1.1\% \\ 
\end{tabular}